---
title: "Plate Reader 19: YPD pH 4, 5.5, 7, 8.5 and without HEPES"
author: "Liz Hughes"
date: "26-09-21"
output:
  html_document:
    toc: yes
    toc_depth: 2
  pdf_document:
    toc: yes
    toc_depth: '2'
  word_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE, message = FALSE, echo = TRUE, cache = FALSE,
  results = "show"
)
```

#Plate reader assay.

Set up to test growth of WT, Gat201 Madhani deletion and Gat201 complemented strains (C26 and C30) in YPD (NA) and YPD + 20mM HEPES at pH= 4 (4), pH=5.5 (5.5), pH = 7 (7) and pH = 8.5 (8.5) at 37 degrees.
3 Biorep each: 3 WT Strains (KN99-alpha), Madhani Gat201 deletion mutant and Gat201-complemented strains 26and 30.
Grow 5ml culture from colony (1 colony = 1 Biorep)in YPD, 30C, 200 rpm ON.
Seed at OD 600nm = 0.2; 200 ul per well.
Run for 432 cycles.



```{r load libraries}
library("reshape2")
library(dplyr)
library(ggplot2)
library(tidyverse)
library(markdown)
library(RColorBrewer)


```



# Read in transposed data as csv file


```{r Load transposed data}
rawdata <- read.csv("../Input/20210926_YPD_pH_TRS.csv")

#view(rawdata)
```

# Change time in seconds to time in hours

```{r Mutate time to hours}

 rawdata_hrs<-mutate(rawdata, Time = Time/3600)

```

# Tidy the data using the melt function from reshape2

```{r Tidy data}
reshaped <- melt(rawdata_hrs, id=c("Time", "Temp"), variable.name="Well",
                 value.name="OD595")


summary(reshaped)
```

# Read in the Plate map data from csv file

```{r read in platemap file}

platemap <- read.csv("../Input/20210926_PR19_YPD_pH_Setup.csv")
head(platemap, n=10)


```


# Combine the reshaped data with the plate map, pairing them by Well

```{r combine the data file and platemap file}
annotated <- inner_join(reshaped, platemap, by="Well")
 
#View(annotated)

```


#Plot all ODs unnormalised

## Do this as check on the data by strain and medium

```{r plot_all_ODs}

ggplot(data=annotated, aes(x=Time, y=OD595, color=Strain, group = Well)) + 
       geom_line() + 
       theme_bw()+
       facet_grid(pH ~ .) +
       labs(x="Time (hrs)", y="Absorbance at 595 nm")
```

#Plot all ODs by Medium, unnormalised

## This allows you to check the data by pH

```{r plot_all_ODs_bypH}

ggplot(data=annotated, aes(x=Time, y=OD595, color=pH, group=Well)) + 
       geom_line() + 
       theme_bw()+
       facet_wrap(Strain ~ .) +
       labs(x="Time (hrs)", y="Absorbance at 595 nm")
```

# Which wells grow to higher OD?

## This helps you to pinpoint any outliers or odd wells with a much higher OD.

```{r find_wells_highOD}

annotated %>%
  group_by(Well) %>% 
  summarise(maxOD=max(OD595)) %>%
  arrange(desc(maxOD))
```



# How stable are ODs of blank wells, to use for normalisation?

```{r plot_blank_ODs}

ggplot(data=filter(annotated,Strain=="", pH=="8.5"), 
       aes(x=Time, y=OD595, color=Strain, group = Well)) + 
       geom_line() + 
       theme_bw()+
       labs(x="Time (hrs)", y="Absorbance at 595 nm")
```

These are fairly very stable.
Check for individual wells that look strange.

```{r plot_individual_blank_wells}

ggplot(data=filter(annotated,Well=="H9"), 
       aes(x=Time, y=OD595, color=Strain, group = Well)) + 
       geom_line() + 
       theme_bw()+
       labs(x="Time (hrs)", y="Absorbance at 595 nm")

```
Remove well  H1 from the analysis.
Remove well  F12 from the analysis.
Remove well  H9 from the analysis.


# Calculate median OD for blank wells to use to normalise data.


```{r calc_blank_medians}
blank_OD_summary <- annotated %>%
  filter(Strain=="") %>%
  filter(Well!= "H1") %>%
  filter(Well!= "F12") %>%
  filter(Well!= "H9") %>%
  group_by(pH) %>%
  summarise(OD_median=median(OD595),
            OD_mean=mean(OD595),
            OD_max=max(OD595),
            OD_min=min(OD595))
print(blank_OD_summary)
```

# Subtract blank OD to make corrected OD and Plot OD_corrected v Time (hrs)
## Normalize data

```{r subtract_blank_OD by pH}

normalisedOD <- annotated %>%
  left_join(blank_OD_summary, by="pH") %>%
  mutate(OD_corrected = OD595 - OD_median)

#view(normalisedOD)
```



#Plot all OD corrected (or normalized)

```{r plot_all_OD_corrected}

ggplot(data=filter(normalisedOD,Strain !=""),  aes(x=Time, y=OD_corrected, color=Strain, group = Well)) + 
       geom_line() + 
       facet_grid(pH ~ .) +
       theme_bw()+
       labs(x = "Time(Hrs)") +
       labs(y = "Absorbance 595nm")
```



# Plot for All strains

```{r plot_all_OD_corrected for all wrap}

ggplot(data=normalisedOD %>%
         filter(Strain != ""),
       aes(Time, OD_corrected, colour = Strain, group = Well)) + 
       geom_line() + 
       geom_hline(aes(yintercept=0.5), linetype="dotted", colour="black") +
      
       theme_bw() +
        scale_y_continuous(limits=c(0,0.5),expand=c(0,0)) +
        labs(x = "Time(Hrs)") +
        labs(y = "Absorbance 595nm") +
      theme(text = element_text(size = 10))

```

```{r select strains}
strain <- filter(normalisedOD, Strain == "KN99a" )

strain2 <-  filter(normalisedOD, Strain == "Gat201-M" )

strain_all <- bind_rows(strain, strain2) 

summary(strain_all)

```


```{r plot stain_all}

ggplot(data=strain_all ,
       aes(Time, OD_corrected, colour = Strain, group = Well)) + 
       geom_line() + 
       geom_hline(aes(yintercept=0.55), linetype="dotted", colour="black") +
   geom_hline(aes(yintercept=0.01), linetype="dotted", colour="black") +
      facet_wrap(pH ~ .)+
       theme_bw() +
        scale_y_continuous(limits=c(0,0.75),expand=c(0,0)) +
        
        labs(x = "Time(Hrs)") +
        labs(y = "Absorbance 595nm") +
        labs(title = "Growth in YPD at different pH")+
      theme(text = element_text(size = 20), line = element_line(brewer.pal(2, "Set1")))
```




# Plot for individual strains

```{r plot_all_OD_corrected KN99a}

ggplot(data=normalisedOD %>%
         filter(Strain == "KN99a"),
         aes(Time, OD_corrected, group = Well, colour = as.factor (pH))) + 
         geom_line() +
         geom_hline(aes(yintercept=0.1), linetype="dotted", colour="black") +
          geom_hline(aes(yintercept=0.2), linetype="dotted", colour="black") +
         theme_bw() +
         scale_y_continuous(limits=c(0,0.75),expand=c(0,0)) +
         labs(x = "Time(Hrs)") +
         labs(y = "Absorbance 595nm") +
         labs(title = "KN99alpha") +
         theme(text = element_text(size = 10))+
         scale_color_brewer(palette="Set1")

   
```

```{r plot_all_OD_corrected Gat201-M}

ggplot(data=normalisedOD %>%
         filter(Strain == "Gat201-M"),
         aes(Time, OD_corrected, group = Well, colour = as.factor (pH))) + 
         geom_line() +
         geom_hline(aes(yintercept=0.1), linetype="dotted", colour="black") +
          geom_hline(aes(yintercept=0.2), linetype="dotted", colour="black") +
         theme_bw() +
         scale_y_continuous(limits=c(0,0.75),expand=c(0,0)) +
         labs(x = "Time(Hrs)") +
         labs(y = "Absorbance 595nm") +
         labs(title = "Gat201-M") +
         theme(text = element_text(size = 10))+
         scale_color_brewer(palette="Set1")

   
```

```{r plot_all_OD_corrected C26}

ggplot(data=normalisedOD %>%
         filter(Strain == "C26"),
         aes(Time, OD_corrected, group = Well, colour = as.factor (pH))) + 
         geom_line() +
         geom_hline(aes(yintercept=0.1), linetype="dotted", colour="black") +
          geom_hline(aes(yintercept=0.2), linetype="dotted", colour="black") +
         theme_bw() +
         scale_y_continuous(limits=c(0,0.75),expand=c(0,0)) +
         labs(x = "Time(Hrs)") +
         labs(y = "Absorbance 595nm") +
         labs(title = "C26") +
         theme(text = element_text(size = 10))+
         scale_color_brewer(palette="Set1")

   
```

```{r plot_all_OD_corrected C30}

ggplot(data=normalisedOD %>%
         filter(Strain == "C30"),
         aes(Time, OD_corrected, group = Well, colour = as.factor (pH))) + 
         geom_line() +
         geom_hline(aes(yintercept=0.1), linetype="dotted", colour="black") +
          geom_hline(aes(yintercept=0.2), linetype="dotted", colour="black") +
         theme_bw() +
         scale_y_continuous(limits=c(0,0.75),expand=c(0,0)) +
         labs(x = "Time(Hrs)") +
         labs(y = "Absorbance 595nm") +
         labs(title = "C30") +
         theme(text = element_text(size = 10))+
         scale_color_brewer(palette="Set1")

   
```